Title

### 2016-06-09 paper/sections/CaseStudy.tex
More case study

>>>>

\subsection{Scans}
% statistics, pictures, implementation
The first case study re-implements \dsl, which is discussed quite thoroughly in
Section~\ref{}.
We strictly follows the flow of the paper and introduce all kinds of
interpretations modularly.
Starting from a plain AST without any interpretations in the base family, we gradually add
\texttt{depth}, \texttt{width}, \texttt{wellSized}, \texttt{layout} and
\texttt{tlayout} methods through extending the family.

The following shows the client code:
\begin{lstlisting}
Circuit circuit =
  above(beside(fan(2), fan(2)),
  above(stretch(fan(2),2,2),
  beside(id(1), beside(fan(2), id(1)))));
\end{lstlisting}

To simplify client code, we defined wrappers around \texttt{of} methods
for conveniently constructing circuits. For instance, the wrapper for
\texttt{Fan} is defined as follows:
\begin{lstlisting}
static Fan fan(int n) {
  return Fan.of(n);
}
\end{lstlisting}
The \texttt{circuit} object supports all the methods stated above.
Additionally, we introduce a \texttt{draw} method which renders a
circuit using Java Swing.
Calling \texttt{circuit.draw()} on the \texttt{circuit} will
display what Figure~\ref{sec:} shows.

% show Java Swing figure
\begin{figure}
\end{figure}

\input{sections/CaseStudyPPrinter.tex}

\subsection{Diagrams}
\emph{diagrams}~\cite{yates2015diagrams} is a declarative Haskell EDSL for creating vector graphics.
This case study implements a simplified version of \emph{diagrams} from
Gibbons's lecture notes~\cite{gibbonsexercises}.

\subsubsection{The Language}
The DSL consists of three simpler sub-languages for describing shapes, styles
and pictures. Note that all these definitions are put inside a base family.

% Haskell code
\begin{comment}
\begin{lstlisting}[language=haskell]
data Shape
  = Rectangle Double Double
  | Ellipse Double Double
  | Triangle Double

type StyleSheet = [Styling]
data Styling
  = FillColour Col
  | StrokeColour Col
  | StrokeWidth Double

data Col = Red | Blue | Bisque | Black | Green | Yellow | Brown

data Picture
  = Place StyleSheet Shape
  | Above Picture Picture
  | Beside Picture Picture
\end{lstlisting}
\end{comment}

\paragraph{Shapes}
Shapes are fundamental components of a picture. The sub-language provides three
primitives for constructing rectangles, ellipses and triangles:

\lstinputlisting[linerange=111-120]{../src/diagrams/Diagrams.java}%APPLY:SHAPE

\paragraph{Styles}
A shape can be decorated with drawing styles. A \texttt{StyleSheet} contains a
list of \texttt{Stylings} for specifying fill color, stroke color or stroke width:

\lstinputlisting[linerange=124-136]{../src/diagrams/Diagrams.java}%APPLY:STYLE
\lstinputlisting[linerange=140-145]{../src/diagrams/Diagrams.java}%APPLY:COLOR

The definition of \texttt{StyleSheet} reveals how to deal with lists or more generally
collection types in \name. We capture the element type using a wildcard with an upper
bound (\texttt{? extends Styling}) to model covariant lists. This allows the
upper bound to be refined with its subtype in extensions. As a result, we are
able to invoke new interpretations on elements of the list.

\paragraph{Pictures}
Provided with a \texttt{StyleSheet}, a \texttt{Shape} can be lifted to a
\texttt{Picture} using \texttt{Place}. Meanwhile, two pictures can be
combined together vertically or horizontally via binary combinator
\texttt{Above} and \texttt{Beside} respectively:

\lstinputlisting[linerange=151-160]{../src/diagrams/Diagrams.java}%APPLY:PICTURE

We can already draw simple pictures based on existing definitions.
For example, Figure~\ref{woman} (with auxiliary wrappers defined) shows
a woman in a red dress and blue stockings.
To actually visualize the picture as shown in the right hand side of
Figure~\ref{woman}, we give semantics to the language for accomplishing this goal.

\begin{figure*}
  \begin{tabular}{lr}
\begin{minipage}[t]{.9\textwidth}
\lstinputlisting[linerange=530-537]{../src/diagrams/Diagrams.java}%APPLY:WOMAN
\end{minipage}
&
\begin{minipage}[t]{.1\textwidth}
  \vspace{0.2em}
\includegraphics[width=1.3cm]{woman.png}
\end{minipage} \\
\end{tabular}
\caption{A woman drawn in \emph{diagrams}}
\label{woman}
\end{figure*}

\subsubsection{Animation}
We use scalable vector graphics (SVG) as the backend for rendering
pictures.
SVG aligns styled shapes by center, which requires us to determine a center
point of a picture as the origin of coordinate, and calculate
positions of all the shapes inside the picture accordingly.
To represent a position in the coordinates we define \texttt{Pos}:
\begin{lstlisting}
@Obj interface Pos {
    double _x(); double _y();
}
\end{lstlisting}
The position of a shape can then be determined by a pair of \texttt{Pos}, called
an extent,
locating the lower left and upper right conners:
\begin{lstlisting}
@Obj interface Extent {
    Pos _p1(); Pos _p2();
}
\end{lstlisting}
Their definitions are stable so we put them out of the family and annotated
with \texttt{@Obj}.

Given an individual shape, its extent can be calculated according to its type.
We therefore extend \texttt{Shape} with a \texttt{toExtent} method:

\begin{lstlisting}
    interface Shape {
        Extent toExtent();
    }
\end{lstlisting}

When pictures are combined together, the center point is changed and hence the
positions of shapes in both pictures should all be recalculated.
The transformation sub-language records how to adjust those positions,
which includes identity transformation, translations, and compositions of the two:

\lstinputlisting[linerange=300-319]{../src/diagrams/Diagrams.java}%APPLY:TRANSFORM

\paragraph{Drawing}
To easily translate a \texttt{Picture} into SVG format, we flatten the recursively structured pictures into a non-empty
sequence of transformed styled shapes.
\texttt{Drawing} contains three lists of the same length carrying all the
information needed for generating a SVG file:

\begin{lstlisting}
interface Drawing {
    List<? extends Shape> _shapes();
    List<? extends StyleSheet> _styles();
    List<? extends Transform> _transforms();
}
\end{lstlisting}

% Step 1: Picture to Drawing
To generate a \texttt{Drawing} from a \texttt{Picture}, we extend the
\texttt{Picture} hierarchy with a \texttt{draw} method:
\lstinputlisting[linerange=268-270]{../src/diagrams/Diagrams.java}%APPLY:DRAW
For Place,

% The implementation for each case is omitted here for space reasons.

% Step 2: Drawing to XML
SVG is based on XML and the structure of an XML file can be captured by following definitions:
\begin{lstlisting}
@Obj interface XML {
    String _tag();
    List<Attr> _attrs();
    List<XML> _xmls();
}
@Obj interface Attr {
    String _name();
    String _value();
}
\end{lstlisting}
We then define \texttt{toXML} method both in \texttt{Drawing} and \texttt{Shape}.
\texttt{Drawing} builds up a complete XML by calling the \texttt{toXML} on each
\texttt{Shape}.

The final step is to pretty print the content of an \texttt{XML} object and
write it to an file with extension
\emph{.svg}.
Then open the file using a modern browser, you will see the right side of
Figure~\ref{woman} appear on the screen.

\subsubsection{Discussion}
\cite{} contains exercises about extensions on the language, e.g.
adding new shapes. The deep embedding approach in Haskell requires modifications on the ADTs
whereas our approach can modularly introduce new language constructs.
Each sub-language retains independent extensibility.
For example, we can add a \texttt{show} method for the purpose of debugging.

### 2016-06-09 paper/sections/Overview.tex
More case study

% introduce of method
As classes are now interfaces, how can we create objects on them?
\name automatically generates a \emph{static} \texttt{of} method for each
class-like interface, i.e. an interface that contains only getters, for creating objects.
For instance, the \texttt{of} method for \texttt{Beside} is:
\begin{lstlisting}
static Beside of(Circuit c1, Circuit c2) {
  return new Beside() {
    Circuit _c1 = c1; Circuit _c2 = c2;
    public Circuit _c1() { return c1; }
    public Circuit _c2() { return c2; }
  }
}
\end{lstlisting}
Inside the \texttt{of} method, an anonymous class that implements
\texttt{Beside} is created and its instance is returned.

@@

Here comes the solution from Gibbons and Wu.
\lstinputlisting[language=haskell,linerange=2-8]{./code/NewSemantics.hs}%APPLY:SEMANTICS_HS
where tuples are used to aggregate multiple interpretations in one definition.
The way to introduce \texttt{desugar} in previous section is similar to this
approach, where we use records instead.
However, neither tuples nor records are extensible - whenever we need a new
interpretation, we have to revise the original code, either expanding the arith of the
tuple or inserting a new field in the record.

@@

Note that we do not explicitly write down \texttt{extends} clauses in the
extended family.
Instead, they are inferred by \name according to their
names and the inheritance relationship between families.
After instrumentation done by \name, the definitions of \texttt{Circuit} and \texttt{Beside}
would be:

\begin{lstlisting}
interface Circuit extends Init.Circuit {
  ... // wellSized declaration
}
interface Beside extends Circuit, Init.Beside {
  Circuit _c1(); Circuit _c2(); // type refinements
  ... // wellSized implementation
}
\end{lstlisting}
Interfaces from base families of the same name will be put after \texttt{extends} clause and the
interface hierarchy of \texttt{Circuit} is preserved in the extended family.
Moreover, \name re-declare getters in the extended \texttt{Beside} so that their return types are refined to be the new \texttt{Circuit}.

@@

Here, we show yet another appealing feature of \name - interpretation combination.
The \texttt{Dependent} family combines two base families

merge interpretations defined for the same data structure together.
%Since \texttt{Circuit} in \texttt{Dependent} family extends the one from
%\texttt{Init}, we can simply call \texttt{width} method on a
%\texttt{Circuit} object when the width of a circuit is needed in the definition
%of \texttt{wellSized}.

@@

\lstinputlisting[language=haskell,linerange=40-41]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES
\emph{stretch ns c} inserts additional wires into the circuit \emph{c} by
summing up \emph{ns} and $above c_1 c_2$ combines two circuits of the same width vertically.
With these new constructs, more complex circuits can be constructed.
For example, Figure~\ref{} shows the circuit constructed by \emph{stretch [2,2,2] (fan 3) `beside` fan 1}.

>>>>

% introduce of method
As classes are now interfaces, how can we create objects on them?
\name automatically generates a \emph{static} \texttt{of} method for each
class-like interface, i.e. an interface that contains only getters, for creating objects.
For instance, the \texttt{of} method for \texttt{Beside} is:
\begin{lstlisting}
static Beside of(Circuit c1, Circuit c2) {
  return new Beside() {
    Circuit _c1 = c1; Circuit _c2 = c2;
    public Circuit _c1() { return _c1; }
    public Circuit _c2() { return _c2; }
  }
}
\end{lstlisting}
Inside the \texttt{of} method, an anonymous class that implements
\texttt{Beside} is created and its instance is returned.

@@

Here comes the solution from Gibbons and Wu:
\lstinputlisting[language=haskell,linerange=2-8]{./code/NewSemantics.hs}%APPLY:SEMANTICS_HS
where tuples are used to aggregate multiple interpretations in one definition.
The way to introduce \texttt{desugar} in Section~\ref{sec:oo} is similar to this
approach, where we use records instead.
However, neither tuples nor records are extensible - whenever we need a new
interpretation, we have to revise the original code, either expanding the arith of the
tuple or inserting a new field in the record.

@@

Note that we do not explicitly write down \texttt{extends} clauses for known members in the
extended family.
Instead, \texttt{extends} clauses are inferred by \name according to their
names and the inheritance relationship stated between families.
After instrumentation done by \name, the definitions of \texttt{Circuit} and \texttt{Beside}
would be:

\begin{lstlisting}
interface Circuit extends Init.Circuit {
  ... // wellSized declaration
}
interface Beside extends Circuit, Init.Beside {
  Circuit _c1(); Circuit _c2(); // type refinements
  ... // wellSized implementation
}
\end{lstlisting}
The corresponding definitions from \texttt{Init} are inherited. For class-like
interfaces like \texttt{Beside}, the new \texttt{Circuit} is also extended.
Moreover, \name re-declare its getters so that their return types are refined to be the new \texttt{Circuit}.

@@

Here, we show yet another appealing feature of \name - interpretation combination.
That is, we can modularly merge interpretations defined for the same data structure together.
The \texttt{Dependent} family combines two base families

%Since \texttt{Circuit} in \texttt{Dependent} family extends the one from
%\texttt{Init}, we can simply call \texttt{width} method on a
%\texttt{Circuit} object when the width of a circuit is needed in the definition
%of \texttt{wellSized}.

@@

\lstinputlisting[language=haskell,linerange=40-41]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES
\emph{stretch ns c} inserts additional wires into the circuit \emph{c} by
summing up \emph{ns} and $above c_1 c_2$ combines two circuits of the same width vertically.
With these new constructs, more complex circuits can be constructed.
For example, Figure~\ref{sec:oo} shows the circuit constructed by \emph{stretch [2,2,2] (fan 3) `beside` fan 1}.

### 2016-06-09 paper/sections/ShallowOO.tex
More case study

\section{Shallow Object-Oriented Programming}

@@

Show the correponding Java programs and the Java program 
with transformation that we can port back to Haskell.

Shallow embedding interpreter pattern
Deep embedding visitor pattern
\end{comment}

In this section, we argue that \emph{shallow embedding} and straightforward OO
programming are closely related. To demonstrate it, we use a subset
of the DSL presented in Gibbons and Wu's paper~\cite{} as the running example.
We first give the original implementation of the language in shallow embedding
and then gradually rewrite the definition into an ``OO style''. Also, we show how to do
transformations in shallow embedding, which is typically thought to be the
strength of deep embedding.

@@

The rewrite goes on by using a record with a sole field to represent the result:
\lstinputlisting[language=haskell,linerange=31-36]{./code/shallowCircuit.hs}%APPLY:CIRCUIT3
We no longer need a separate \texttt{width3} function as the field name is the extractor.
The implementation is still shallow since essentially it is nothing different
from the last one.
However, the definition already approaches an OO program such that we can easily
translate it into an OO language like Java:
\lstinputlisting[linerange=7-26]{../src/paper/sec2/Circuit.java}%APPLY:CIRCUIT_OO
where the datatype maps to the abstract class \texttt{Circuit} with its field declared as an abstract method.
Each case in the semantic function corresponds to an class and its parameters become fields of the class.
And all the classes extends \texttt{Circuit} by giving \texttt{width} a definition.

@@

An instance of this kind of operations could be desugaring, which eliminates some
language constructs that could be expressed by other language constructs without
sacrificing the expressiveness of the language.
In \dsl, the \emph{identity n} construct is such a syntactic sugar that can
be rewritten via the following formula:
$$identity\ n = (fan\ 1) \overbrace{\ beside\ (fan\ 1)}^{n-1}\ (n > 0)$$
which states that \emph{identity n} can be represented as \emph{n} of \emph{fan 1} combined with \emph{beside}.
%Using the above formula, we can define a transformation that desugars every
%occurrence of \emph{identity} to a combination of \emph{fan} and \emph{beside}.

>>>>

\section{Shallow Object-Oriented Programming}\label{sec:oo}

@@

Show the correponding Java programs and the Java program 
with transformation that we can port back to Haskell.
\end{comment}

In this section, we argue that \emph{shallow embedding} and straightforward OO
programming are closely related. To demonstrate it, we use a subset
of the DSL presented in Gibbons and Wu's paper~\cite{} as the running example.
We first give the original implementation of the language in shallow embedding
and then gradually rewrite the definition into an ``OO style''. Also, we show how to do
transformations in shallow embedding, which is typically thought to be the
privilege of deep embedding.

@@

The rewrite goes on by using a record with a sole field to represent the result:
\lstinputlisting[language=haskell,linerange=31-36]{./code/shallowCircuit.hs}%APPLY:CIRCUIT3
We no longer need a separate \texttt{width3} function as the field name is the extractor.
The implementation is still shallow since essentially it is nothing different
from the last one.
However, the definition already approaches an OO program such that we can easily
translate it into an OO language like Java:
\lstinputlisting[linerange=7-26]{../src/paper/sec2/Circuit.java}%APPLY:CIRCUIT_OO
where the record type maps to the abstract class \texttt{Circuit} and field
declaration becomes method declaration.
Each case in the semantic function corresponds to an class and its parameters become fields of the class.
And all the classes extends \texttt{Circuit} by giving \texttt{width} a definition.

@@

An instance of this kind of operations could be desugaring, which eliminates some
language constructs that could be expressed by other language constructs without
sacrificing the expressiveness of the language.
In \dsl, the \emph{identity n} construct is such a syntactic sugar that can
be rewritten via the following formula:
$$identity\ n = (fan\ 1) \overbrace{\ beside\ (fan\ 1)}^{\text{repeat }n-1\text{ times}}\ (n > 0)$$
which states that \emph{identity n} can be represented as \emph{n} of \emph{fan 1} combined with \emph{beside}.
%Using the above formula, we can define a transformation that desugars every
%occurrence of \emph{identity} to a combination of \emph{fan} and \emph{beside}.
